---
title: "Parent Child Height Modeling and Prediction Application Presentation"
author: "Helena Luk"
date: "January 22, 2022"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(UsingR)
data("galton")
```

## Background

This application is for modeling and prediction of child's height using Parents' Data.  It is made interactive using Shiny.  The model is created using the tabulated data set that Galton used in 1885 to study the relationship between a parent's height and their children. 

In the data set, there are 928 observations that contain 205 different parents and 928 children with the following 2 variables:  
- Child:  The child's height  
- Parent: The "midparent's" height  

The midparent's height is an average of the father's height and 1.08 times the mother's.

## Data Sample 
In this web application, the users could select a sample of data for modeling by dragging a box around the black dots in the following graph.  

```{r dataplot}
plot(galton$parent, galton$child, xlab = "Parent (in inches)",
      ylab = "Child (in inches)", main = "Galton's Heights Data of Parents & their Children", cex = 1.0, pch = 16, bty = "n")
```

## Resulting Model

A linear regression model will then be generated to find the intercept and the slope.  For example, if the whole Galton data set is used, the following r code will generate the model:

```{r model, echo=TRUE}
  model <- lm(child ~ parent, data = galton)
  intercept <- model$coefficients[1]
  slope <- model$coefficients[2]
```

The resulting intercept will be `r intercept` with slope `r slope`.

## Prediction

Once the model is generated using the sample of data selected by the user, the user could then slide the bar to select the "midparent's" height.  The resulting prediction of the child's height will then be displayed at the bottom.  

```{r prediction, echo=TRUE}
  childheight <- intercept + slope * 65
```

For example, if the whole Galton data set is used for generating the model as in previous slide, we could use the above r code to predict the child's height for midparent's height of 65 inches. The resulting height of the child is `r childheight` inches.

By varying the midparent's height via sliding the bar, it will result in different child's height.  On the other hand, a different model can be selected again with the midparent's height unchanged.  With the model changed, a different child's height will result.








